Open-source landscape for 3D CSEM modelling

### 2020-10-02 manuscript/Manuscript.tex

\begin{abstract}
% The (points) are from Mahrer (1993), Geophysics, Some Considerations When Abstracting; DOI: 10.1190/1.58010001.1
% (1) principal objectives and scope of the work
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. However, this has changed over the last few years, and today there exists a selection of different open-source codes which are up to the task. We have a closer look at four different ones, where in all cases at least the user-facing functionality is written in Python.
% (2) methodology
For the numerical comparison we use three models: a simple layered model; a model containing three resistive blocks in a layered background; and a complex, real-world resistivity model. All three models are vertically transverse isotropic. The analysis includes three different mesh-types, simple tensor meshes, octree meshes, and unstructured tetrahedral meshes. Two of the codes use finite volumes and the other two use finite elements.
% (3) results
The comparison shows that all four codes are able to compute the different models to sufficient precision. The biggest difference between codes is actually the discretization and the ease of creating a suitable mesh, particularly as that part often takes much more time than the actual computation. The required runtime and memory varies greatly between the codes. However, these depend mainly on the underlying solver, and not on the actual CSEM codes.
% (4) conclusions
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community at large, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. All our scripts to reproduce the shown results are open-source.
%
\\~\\ \# \# Alternative \# \# \\
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. This has changed over the last few years, and today there exists a selection of different open-source codes. Using four different codes in the Python ecosystem we model responses for three increasingly complex models. The verification of the results with semi-analytical solutions for a simple case as well as the validation between the codes and with results from the industry for more complex models show that the open-source codes are able to compute CSEM responses for challenging, large-scale models. Our comparison includes finite-element and finite-volume codes, using structured tensor and octree meshes as well as unstructured tetrahedra meshes. Our results show further that accurate responses can be obtained independently of the chosen method and the chosen mesh. The difficult part, however, is to appropriately discretize the model. The runtime and memory requirements, however, vary greatly between different codes. But we have found that much more time was spent on creating the appropriate input files than running the actual computation. We provide the three models and the corresponding responses of four codes, which can be used for verification of new and existing codes. The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community at large, and we invite and encourage the community to make more code, modelling scripts, and results publicly available.
\end{abstract}

@@

    \item {\color{red}\bfseries This version is quite different from the previous you revised. However, from now on we do not expect too many changes. To make it easier for everyone to see the new changes please make sure that from now on "Track Changes" is always on in Overleaf.}

@@

    \item We probably should be careful to not mention the actual code names too often. E.g., I removed them from the abstract.
    \item I tried to avoid double reference. E.g., SciPy, MUMPS, pyGIMLi: I referenced them the first time they are mentioned, and not afterwards. Current exceptions: Miensopust et al., 2013, is cited twice.

@@

            \item Kerry Key (domain knowledge with \texttt{MARE3DEM}, also experience with open-sourcing with \texttt{MARE2DEM});

@@

            \item Dimitry Avdeev and Ralph-Uwe Börner (they both did extensive 3D code overview/review papers in 2005 and 2010, respectively, and it would be interesting to get their feedback).

@@

@@

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most common applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or octree meshes \citep{ECP.07.Haber}.

Very well written overviews about the different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and \cite{SG.10.Borner}. But there was also a tremendous publications output with regards to 3D EM modelling in the last 10-15 years, at least partly driven by the ever increasing computing power. One reason why there are so many publications about this topic results from the variety of techniques to solve the systems of linear equations (SLE). They can be distinguished between direct solvers \citep{GEO.09.Streich, GP.14.Chung, GEO.14.Jaysaval, GEO.15.Grayver, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GJI.16.Jaysaval}.

@@

It is becoming more common for researchers in many domains of science to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an \emph{open model of development} where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of (open-source) code towards building global communities around open projects. Other notable Python projects with an open model of development within the same realm as the codes under considerations are \texttt{pyGIMLi} \citep{CAG.17.Rucker}, \texttt{Fatiando} \citep{JOSS.18.Uieda}, \texttt{GemPy} \citep{GMD.19.DeLaVarga}, and \texttt{PyVista} \citep{JOSS.19.Sullivan}.

We discuss the topic of open-source developments with respect to 3D CSEM modelling. Even though it comprises still the minority of developments, the open-source landscape in this field of research changed in the last five years quite dramatically. Previously, there existed only closed-source codes owned by companies and consortia, or individual projects that were only available directly from the authors upon request. These codes often came without proper documentation. But more importantly, they also often came without a proper license, which essentially means that they are not open source (no license means legally speaking that, in most countries of the world, all rights are reserved by the author; see, for instance, the \emph{Berne Convention}). We introduce and compare four recent open-source projects with a focus on 3D CSEM modelling. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes.

Furthermore, we deal with the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. This is, to the best of our knowledge, the first comparison exercise for CSEM codes. We hope our efforts help to ensure the reliability of simulations for any new or existing, open-source or closed-source 3D CSEM project.

@@

\itodo{Ones were ready check if we should introduce Maxwell's equations or not. Miensopust et al., 2013, for instance, do not touch on any equation.}

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{NAT.20.Harris} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them are not just open-source but follow the open model of development, meaning that they come with both, an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as \texttt{conda}, \texttt{docker}, or \texttt{pip} significantly supported us to make our codes easily available and installable on any of the major operating system without any user-side compilations. Each one can be downloaded and installed with a \mycom[DW]{I guess this statement is currently a bit of a stretch for \petgem, correct Octavio?}single command.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the different codes is given in Table~\ref{tbl:codecomparison}. All codes have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range commonly applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator.
\imycom[DW]{I took out «the weak formulation of», as it actually doesn't apply to \emg3d. Please make sure that everything stated above («have in common that...») actually applies to your code.}

A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations PEC and PMC, which stand for perfect electric conductor and perfect magnetic conductor, respectively. Another abbreviation used in the tables is dof for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation or CPU time is therefore much higher than the reported runtime for the codes that run in parallel. Memory refers to the maximum memory increase at any point of solving the SLE.

% Lindsey, I leave the SciPy solvers out. You mention that in practice they are not often used. emg3d can also use those solvers, and then multigrid is used as preconditioner. However, we cite SciPy, and as you say, SimPEG mostly uses PARDISO or MUMPS. To keep it simply I think we can leave SciPy out on the solver side. I mention it in the emg3d-section, if you want you can mention it in the SimPEG section too.
\tabl[btp]{codecomparison}{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of hardware, software, and operator with which the different codes were run.\itodo{Could everyone check if their RAM is GB or GiB?}}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 (server)}                  & Raphael Rochlitz \\
          & - 144 Xeon Gold 6154 CPU @2.666 GHz               & \\
          & - $\approx3$ TB DDR4 RAM                          & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \emg3d  & \textbf{Dell Latitude (laptop)}                   & Dieter Werthmüller \\
          & - i7-6600U CPU@2.6 GHz x4                         & \\
          & - 15.5 GiB RAM                                    & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \petgem & \textbf{Marenostrum4 (supercomputer)}             & Octavio Castillo-Reyes \\
          & - Intel Xeon Platinum from Skylake generation; 2 sockets Intel Xeon Platinum 8160 CPU with 24 cores each @2.10GHz for a total of 48 cores per node & \\
          & - 386 GB DDR4 RAM per node                        & \\
          & - SuSE Linux Enterprise                           & \\[.5em]
  \simpeg & \textbf{GKE n2-custom (Google cloud)}             & Lindsey Heagy \\
          & - Intel Cascade Lake, 8 vCPUs                     & \\
          & - 355 GB RAM                                      & \\
          & - Ubuntu 16.04                                    & \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity is mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarisation parameters in frequency-domain and three methods for simulating time-domain responses were added to \custem. The provided meshing tools are based on \texttt{TetGen} \citep{TOM.15.Si} and functionalities of \texttt{pyGIMLi} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

The \custem toolbox is under steady development and maintenance. Future enhancements comprise adding iterative solution techniques, provided by \texttt{FEniCS}, as well as further functionalities for automatically incorporating intersecting bodies such as folds or faults in meshes. Furthermore, current work is on using \custem for inverse modelling applications.

@@

As such the code is, contrary to the other three, primarily a solver, which solves the Maxwell's equation for a single frequency and a single, electric source. However, recent developments added functionalities such that \emg3d can be used directly as a more general EM modeller too. It features now routines to obtain the magnetic field due to an electric source, to obtain the electric and magnetic fields due to a magnetic source, and also to obtain time-domain responses. The most recent development added the possibility to compute the responses for entire surveys, and also to compute the misfit and the gradient of the misfit function, and \emg3d is as such ready to be used in inversion schemes.

For the underlying discretization the Python package \discretize is used, which is part of the larger \simpeg ecosystem. Interoperability between \simpeg and \emg3d is therefore straightforward.

@@

Based on the current state of the \petgem project, there are many possibilities to improve or to add additional features. For instance, in the short-term, more modelling routines and support for multi-source set-ups will be implemented. For the long-term, the obvious next step is to move to 3D CSEM data inversion.

\subsection{SimPEG}

\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarisation, electromagnetics (time and frequency, controlled and natural sources such as magnetotellurics), and fluid flow (Richards equation). It is a community driven project that aims to support researchers and \mycom[DW]{I added «and practitioners», as I think «support researchers» downplays the fact that \simpeg is used by various companies as well\dots}practitioners by providing a flexible, extensible framework and to facilitate integration of data sets from a variety of geophysical methods, including joint inversions \citep{GJI.20.Astic}, by providing a common interface to each method.

@@

Please not that \emg3d and \simpeg used the same tensor mesh in this example.
\imycom[DW]{Lindsey: A sentence why Tensor over OcTree.}
\imycom[DW]{Raphael/Octavio: A sentence about the differences in the meshes of
\custem and \petgem.}

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
\imycom[DW]{Lindsey: A note about high RAM usage of \simpeg? About PARDISO or Mumps or similar. Could be used as a general note that it boils down to the underlying solver.}

%
\tabl[btp]{comp-layered}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 24 &   118 &  97.8 & 1530808 \\ % dof: 3061616 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &    90 &  34.2 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The results for the three receiver lines $y=-3,0,+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the NRMSDs between the absolute responses of the codes.
%
\plot*{results-block}{width=\textwidth}{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the NRMSDs (\%) between the absolute responses ($|E_x|$) of the different codes are shown in the bottom row.}
%
The NRMSD is generally below a few percent, except for large offsets. An interesting feature is that the NRMSD between \emg3d and \simpeg is very low except at the boundary. The reason is that the codes use the same tensor mesh for modelling, but while \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition.
\mycom[DW]{Lindsey, what about the different formulations we use?}
\todo{expand the analysis of the Block Model -- any ideas / inputs?}

>>>>

\begin{abstract}
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. This has changed over the last few years, and today there exists a selection of different open-source codes. Using four different codes in the Python ecosystem we model responses for three increasingly complex models. We first verify the results with semi-analytical solutions for a simple case. Then we validate the responses of more complex models between the different codes and with results from the industry. These validations show that the open-source codes are able to compute CSEM responses for challenging, large-scale models. Our comparison includes finite-element and finite-volume codes using structured tensor and octree meshes as well as unstructured tetrahedra meshes. Our results show further that accurate responses can be obtained independently of the chosen method and the chosen mesh. The runtime and memory requirements vary greatly based on the choice of iterative or direct solvers. But we have found that much more time was spent on creating the appropriate input files than running the actual computation. The challenging task is to appropriately discretize the model, independently of the chosen mesh and the chosen solver. We provide the three models and the corresponding responses of four codes, which can be used for verification of new and existing codes. The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further.
\end{abstract}

@@

@@

@@

            \item Kerry Key (domain knowledge with \texttt{MARE3DEM}, also experience with open-sourcing with \texttt{MARE2DEM}); \imycom[RR]{Agreed}

@@

            \item Dimitry Avdeev and Ralph-Uwe Börner (they both did extensive 3D code overview/review papers in 2005 and 2010, respectively, and it would be interesting to get their feedback). \imycom[RR]{Strongly Agreed}

@@

    \imycom[RR]{Agreed.}

@@

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most commonly applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or octree meshes \citep{ECP.07.Haber}.

Very well written overviews about the different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and \cite{SG.10.Borner}. But there was also a tremendous publications output with regards to 3D EM modelling in the last 10-15 years, at least partly driven by the ever increasing computing power. One reason why there are so many publications about this topic results from the variety of techniques to solve the systems of linear equations (SLE). They can be distinguished between direct solvers \citep{GEO.09.Streich, GP.14.Chung, GEO.14.Jaysaval, GEO.15.Grayver, SEG.15.Oh, GJI.18.Wang}, indirect solvers \citep{GP.06.Mulder, GJI.15.Jaysaval}, or a combination of both, so-called hybrid solvers \citep{GEO.18.Liu}. The solvers often use preconditioners such as the multigrid method \citep{SIAM.02.Aruliah, GJI.16.Jaysaval}.

% The remaining paragraphs should be shortened/condensed

@@

Today it is becoming more common for researchers in many domains of science to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an \emph{open model of development} where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of (open-source) code towards building global communities around open projects. Other notable Python projects with an open model of development within the same realm as the codes under considerations are \texttt{pyGIMLi} \citep{CAG.17.Rucker}, \texttt{Fatiando} \citep{JOSS.18.Uieda}, \texttt{GemPy} \citep{GMD.19.DeLaVarga}, and \texttt{PyVista} \citep{JOSS.19.Sullivan}.

The open-source landscape in this field of research changed in the last five years quite dramatically. We introduce and compare four recent open-source projects with a focus on 3D CSEM modelling. All presented codes are in the Python ecosystem and use either the FV method on structured grids or the FE method on unstructured tetrahedral meshes. As such, we address the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. This is, to the best of our knowledge, the first comparison exercise for CSEM codes. We hope our efforts help to ensure the reliability of simulations for any new or existing, open-source or closed-source 3D CSEM project.

@@

@@

The four codes under consideration are, in alphabetical order, \custem \citep{GEO.19.Rochlitz}, \emg3d \citep{JOSS.19.Werthmuller}, \petgem \citep{CAG.18.CastilloReyes, GJI.19.CastilloReyes}, and \simpeg \citep{CAG.15.Cockett, CAG.17.Heagy}. All four codes have their user-facing routines written in Python, and all of them make heavy use of \texttt{NumPy} \citep{NAT.20.Harris} and \texttt{SciPy} \citep{NM.20.Virtanen}. The four of them are not just open-source but follow the open model of development, meaning that they come with both, an open-source license and an online-hosted version-control system with tracking possibilities (raising issues, filing pull requests). All developments comprise an extensive online documentation with many examples and have continuous integration to some degree. Newer package-management systems such as \texttt{conda}, \texttt{docker}, or \texttt{pip} significantly supported us to make our codes easily available and installable on any of the major operating system without any user-side compilations.

In the following, we provide a quick overview of the codes. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the different codes is given in Table~\ref{tbl:codecomparison}. All codes have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range commonly applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator.
\imycom[DW]{I took out «the weak formulation of», as it actually doesn't apply to \emg3d. Please make sure that everything stated above («have in common that...») actually applies to your code.
\imycom[RR]{everything fits to custEM}
}
A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations PEC and PMC, which stand for perfect electric conductor and perfect magnetic conductor, respectively. Another abbreviation used in the tables is dof for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation or CPU time is therefore much higher than the reported runtime for the codes that run in parallel. Memory refers to the maximum memory increase at any point of solving the SLE.

% Lindsey, I leave the SciPy solvers out. You mention that in practice they are not often used. emg3d can also use those solvers, and then multigrid is used as preconditioner. However, we cite SciPy, and as you say, SimPEG mostly uses PARDISO or MUMPS. To keep it simply I think we can leave SciPy out on the solver side. I mention it in the emg3d-section, if you want you can mention it in the SimPEG section too.
\tabl[bthp]{codecomparison}{Comparison of the four codes under consideration. Note that \emg3d is a solver on its own, while the other codes implement third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar}, \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO} \citep{FGCS.04.Schenk}.}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PEC; PMC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[bthp]{machines}{List of hardware, software, and operator with which the different codes were run.\itodo{Could everyone check if their RAM is GB or GiB?}}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 (server)}                  & Raphael Rochlitz \\
          & - 144 Xeon Gold 6154 CPU @2.666 GHz               & \\
          & - $\approx3$ TB DDR4 RAM                          & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \emg3d  & \textbf{Dell Latitude (laptop)}                   & Dieter Werthmüller \\
          & - i7-6600U CPU@2.6 GHz x4                         & \\
          & - 15.5 GiB RAM                                    & \\
          & - Ubuntu 18.04                                    & \\[.5em]
  \petgem & \textbf{Marenostrum4 (supercomputer)}             & Octavio Castillo-Reyes \\
          & - Intel Xeon Platinum from Skylake generation; 2 sockets Intel Xeon Platinum 8160 CPU with 24 cores each @2.10GHz for a total of 48 cores per node & \\
          & - 386 GB DDR4 RAM per node                        & \\
          & - SuSE Linux Enterprise                           & \\[.5em]
  \simpeg & \textbf{GKE n2-custom (Google cloud)}             & Lindsey Heagy \\
          & - Intel Cascade Lake, 8 vCPUs                     & \\
          & - 355 GB RAM                                      & \\
          & - Ubuntu 16.04                                    & \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

The toolbox considers generally anisotropic petrophysical properties. Even though changes of the conductivity are mainly of interest for CSEM modelling, the electric permittivity and magnetic permeability can be taken into account using the preferred electric field approach on Nédélec elements. Recently, induced polarisation parameters in frequency-domain and three methods for simulating time-domain responses were added to \custem. The provided meshing tools are based on \texttt{TetGen} \citep{TOM.15.Si} and functionalities of \texttt{pyGIMLi} facilitate the generation of tetrahedral meshes including layered-earth geometries with topography or bathymetry and anomalous bodies which are allowed to be connected or to reach the surface.

@@

As such the code is, contrary to the other three, primarily a solver, which solves the Maxwell's equation for a single frequency and a single, electric source. However, recent developments added functionalities such that \emg3d can be used directly as a more general EM modeller too. It features now routines to obtain the magnetic field due to an electric source, to obtain the electric and magnetic fields due to a magnetic source, and also to obtain time-domain responses. For the underlying discretization the Python package \discretize is used, which is part of the larger \simpeg ecosystem. Interoperability between \simpeg and \emg3d is therefore straightforward.

@@

% Octavio: We removed the outlook part, we did the same for custEM and emg3d. I hope that is OK.
% Based on the current state of the \petgem project, there are many possibilities to improve or to add additional features. For instance, in the short-term, more modelling routines and support for multi-source set-ups will be implemented. For the long-term, the obvious next step is to move to 3D CSEM data inversion.

\subsection{SimPEG}
\todo{Please condense SimPEG a bit; the other codes use about 200-240 words, SimPEG currently close to 400.}
\simpeg is a modular toolbox for simulations and gradient based inversions in geophysics. Current functionality includes modelling and inversion capabilities for gravity, magnetics, direct current resistivity, induced polarisation, electromagnetics (time and frequency, controlled and natural sources such as magnetotellurics), and fluid flow (Richards equation). It is a community driven project that aims to support researchers and \mycom[DW]{I added «and practitioners», as I think «support researchers» downplays the fact that \simpeg is used by various companies as well\dots}practitioners by providing a flexible, extensible framework and to facilitate integration of data sets from a variety of geophysical methods, including joint inversions \citep{GJI.20.Astic}, by providing a common interface to each method.

@@

Please not that \emg3d and \simpeg used the same tensor mesh in this example.
\imycom[DW]{Lindsey: A sentence why Tensor was chosen over OcTree here?}

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.

%
\tabl[bthp]{comp-layered}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 24 &   118 &  97.8 & 1530808 \\ % dof: 3061616 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &    90 &  34.2 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The results for the three receiver lines $y=-3,0,+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the NRMSDs between the absolute responses of the codes.
%
\plot*{results-block}{width=\textwidth}{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the NRMSDs (\%) between the absolute responses ($|E_x|$) of the different codes are shown in the bottom row.}
%

There are a few noteworthy points:

\begin{enumerate}
  \item The error levels are overall comparable; the NRMSD is generally below a few percent, increasing towards larger offsets.
  \item The NRMSD between the FV codes \emg3d and \simpeg is very low except at the boundary. The generally very low NRMSD is because these codes use the same rectilinear mesh for modelling. The increasing NRMSD towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition.
  \item The FE codes \custem and petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
\end{enumerate}

The corresponding required runtime and memory are listed in Table~\ref{tbl:comp-block}.
%
\tabl[bthp]{comp-block}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the block model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
\toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 24 &   158 & 115.9 & 1654242 \\ % dof: 3308484 / 2  (Real+Imag -> Complex)
  \emg3d  &  1 &   115 &   0.4 & 4813216 \\
  \petgem & 24 &    72 &  35.8 & 2455868 \\
  \simpeg &  4 & 10295 & 280.4 & 4813216 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

\imycom[RR]{@ Octavio: I think you use your own mesh here? Do you use the iterative solver in this example? otherwise, the values in the table do not make too much sense for me...}
